Deep mutational scanning of hemagglutinin helps predict evolutionary fates of human H3N2 influenza variants

### 2018-06-27 paper/reviewer_response_PNAS.tex

\title{Response to reviews of ``Deep mutational scanning of hemagglutinin helps predict evolutionary fates of human H3N2 influenza variants'' for \textit{PNAS}}
\author{Juhye M. Lee, John Huddleston, Michael B. Doud, Kathryn A. Hooper,\\Nicholas C. Wu, Trevor Bedford, Jesse D. Bloom}

\begin{document}
\maketitle

We thank the reviewers for the careful reading of our manuscript.
We were gratified that they recognized the quality of our work: one of the reviewers said, ``the experimental setup is sound and there is strong correlation among technical and biological replicates.''

The main contribution of our work is to show that deep mutational scanning of an H3 HA provides information that is informative about the evolutionary fates of H3N2 influenza virus variants in nature.
We appreciated the reviewers' recognition of the important step our work takes in linking experiments with quantitative models of influenza virus evolution, and their recognition that we carefully discuss the caveats as well as the strengths of our approach.
Reviewer \#1 did make an important suggestion to look at HA sequences only from unpassaged virus isolates---in the revised version we have done that, and found that our data are even more predictive of evolutionary fate for these higher quality sequences.

@@

We are excited about the future directions that the reviewer suggests.
However, the current study is a complete and important contribution even if (as is almost invariably the case in science) future work can probably take things further.
Researchers have been performing functional experiments on influenza virus for decades, and have been trying to forecast the virus's evolution for decades---yet ours is the first study to successfully use such functional experiments to predict the success of viral strains in nature. 
We have gone to great pains to fairly discuss limitations and caveats and how things could be improved in the future, but that should not detract from the advances in the current paper.
}

\subsubsection*{Minor Points:}

The correlation for prospective analysis (post-Perth) is smaller than the correlation for post-hoc analysis (pre-Perth). Since prospective analysis is more important when predicting evolutionary fates, it might be beneficial for the authors to comment on this difference.

\response{
The reviewer is correct that on the complete data sets in the original submission, the correlation was slightly higher for the post-hoc (post-Perth) and prospective (pre-Perth) analysis.
But at the suggestion of Reviewer \#1, we have now also included a post-Perth analysis using \emph{only} HA sequences from unpassaged viral isolates (there are not enough unpassaged isolates from before 2009 to perform a similar unpassaged pre-Perth analysis).
As shown in the new Figure 5, the correlation for the unpassaged post-Perth is actually higher than either other set, suggesting that there may not be a systematic trend for poorer performance in prospective analyses.
}

### 2018-06-28 paper/manuscript.tex
edited dN/dS text per Trevor suggestions

\subsection*{Our measurements are consistent with existing knowledge about HA's evolution and function}
How do the HA amino-acid preferences measured in our experiments relate to the evolution of H3N2 influenza virus in nature?
This question can be addressed by evaluating how well an experimentally informed codon substitution model (ExpCM) using our measurements describes H3N2 evolution compared to standard substitution models~\cite{bloom2017identification,hilton2017phydms}.
Table~\ref{tab:phydms} shows that an ExpCM using the across-replicate average of our measurements greatly outperforms conventional substitution models.
This result indicates that our experiments authentically capture some of the constraints on HA evolution.
A substitution model in which the amino-acid preferences were averaged across all sites (ExpCM, site avg.) performs no better than conventional substitution models, demonstrating that the reason that our measurements are informative is because they capture site-specific constraints.
The relative rate of nonsynonymous to synonymous substitutions (dN/dS or $\omega$) is $\ll$1 for conventional substitution models (Table~\ref{tab:phydms}), since the only way that these models can account for the constraints on HA evolution is via the heuristic use of the $\omega$ parameter to depress the rates of all nonysynonymous substitutions.
However, the relative rate of nonsynonymous to synonymous substitutions after accounting for the amino-acid preferences measured in our experiments ($\omega$ for the ExpCM) is close to one (Table~\ref{tab:phydms}), indicating that most purifying selection against nonsynonymous substitutions is accounted for in the deep mutational scanning.
The ExpCM stringency parameter~\cite{hilton2017phydms} is 2.47 (Table~\ref{tab:phydms}), indicating that natural selection favors the same amino acids as the experiments but with greater stringency.
Throughout the rest of this paper, we use the amino-acid preferences re-scaled~\cite{bloom2017identification,hilton2017phydms} by this stringency parameter.
These re-scaled preferences are shown in Figure~\ref{fig:logoplot}.

>>>>

\subsection*{Our measurements are consistent with existing knowledge about HA's evolution and function}
How do the HA amino-acid preferences measured in our experiments relate to the evolution of H3N2 influenza virus in nature?
This question can be addressed by evaluating how well an experimentally informed codon substitution model (ExpCM) using our measurements describes H3N2 evolution compared to standard substitution models~\cite{bloom2017identification,hilton2017phydms}.
Table~\ref{tab:phydms} shows that an ExpCM using the across-replicate average of our measurements greatly outperforms conventional substitution models.
This result indicates that our experiments authentically capture some of the constraints on HA evolution.
A substitution model in which the amino-acid preferences were averaged across all sites (ExpCM, site avg.) performs no better than conventional substitution models, demonstrating that the reason that our measurements are informative is because they capture site-specific constraints.
The relative rate of nonsynonymous to synonymous substitutions (dN/dS or $\omega$) is $\ll$1 for conventional substitution models (Table~\ref{tab:phydms}).
However, the relative rate of nonsynonymous to synonymous substitutions after accounting for the amino-acid preferences measured in our experiments ($\omega$ for the ExpCM) is close to one (Table~\ref{tab:phydms}), indicating that most purifying selection against nonsynonymous substitutions is accounted for in the deep mutational scanning.
The ExpCM stringency parameter~\cite{hilton2017phydms} is 2.47 (Table~\ref{tab:phydms}), indicating that natural selection favors the same amino acids as the experiments but with greater stringency.
Throughout the rest of this paper, we use the amino-acid preferences re-scaled~\cite{bloom2017identification,hilton2017phydms} by this stringency parameter.
These re-scaled preferences are shown in Figure~\ref{fig:logoplot}.
